Final draft

### 2021-07-19 nuova final draft.Rmd
l'unico file importante è nuova final draft

---
header-includes:
- \def\code#1{\texttt{#1}}
- \usepackage{amssymb}
- \usepackage{txfonts}
- \usepackage{amsmath}
- \usepackage[bindingoffset=1.5cm, left=3cm, right=3cm, top=3cm, bottom=3cm]{geometry}
- \usepackage{graphicx}
- \usepackage{MnSymbol} %allows to insert the QED at the end of the proofs
- \usepackage{hyperref} %set the hyperlink to cite sections, equations...
- \urlstyle{same}
- \usepackage{comment}
- \usepackage{enumitem}
- \usepackage{ntheorem}
- \usepackage{lipsum}
- \usepackage{xcolor}
- \theoremstyle{break}
- \newtheorem{thm}{Theorem}% theorem counter resets every \subsection
- \renewcommand{\thethm}{\arabic{thm}}
- \newtheorem{proposition}{Proposition}
- \newtheorem{assumption}{Assumption}
- \newtheorem{lemma}{Lemma}
- \newtheorem{definition}{Definition}
- \theoremstyle{nonumberplain}
- \newtheorem{proof*}{Proof.}
- \renewcommand{\P}{\mathbb{P}}
- \newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
- \def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
- \setcounter{MaxMatrixCols}{10}
- \setlength{\oddsidemargin}{.0in}
- \setlength{\textwidth}{6 in}
- \setlength{\topmargin}{-.1in}
- \setlength{\textheight}{8.1in}
- \renewcommand{\baselinestretch}{1.5}
- \newcommand{\E}{\mathbb E}
- \newcommand{\Q}{\mathbb Q}
- \newcommand{\M}{\mathbb M}
- \newcommand{\R}{\mathbb R}
- \newcommand{\B}{\mathcal B}
- \newcommand{\X}{\mathcal X}
- \newcommand{\Y}{\mathcal Y}
- \newcommand{\Pd}{\mathbb P}
- \newcommand{\N}{\mathcal N}
- \newtheorem{algorithm}{Algorithm}[section]
- \usepackage{float}
- \floatplacement{figure}{}

title: "Final draft"

@@

output:
  pdf_document: 
   latex_engine: xelatex
   fig_width: 5
   fig_height: 3.5
   fig_caption: true
   extra_dependencies: ["float"]
   number_sections: true

@@

Let $\varphi: \X \rightarrow \R$ be a measurable, integrable and bounded real-valued function, where $(\X,\B(\X),\Q)$ is a probability space.\footnote{$\B(\X)$ is the Borel sigma-algebra of $\X$.}
Suppose that the probability measure $\Q$ admits a density $q(x)$ with respect to some measure $\nu(dx)$ and it is feasible to sample from that density.\
To estimate the expected value of $\varphi$ with respect to the density $q$, the basic Monte Carlo (MC) method would require to sample $N$ times from the target density $q$ and to compute the average of the function $\varphi$ values. $\hat \varphi_{MC}$ is the standard MC estimator:
\begin{equation*}
   \hat \varphi_{MC}=\frac{1}{N} \sum_{n=1}^N\varphi (X^n)\approx \E_q(\varphi)~~~X^n\thicksim q(\cdot)
\end{equation*}
However, in many circumstances either it is impossible to sample from the target density or there exist another density - call it $m(\cdot)$ - different from $q$ that produces ``more efficient" estimates. 
In all the cases for which sampling takes place from a distribution different from the ``correct" (i.e., the true) one, importance sampling and resampling methods can be applied (as augmented versions of the standard MC algorithm).

@@

Let us consider a more general state-space model, where the state vector includes a time-constant parameter vector $\psi \in \Psi$, where $\Psi$ is the parameter space. Since we can interpret the parameter as a state with the law of motion $\psi_t=\psi_{t-1}=\psi$, then it is possible to adopt a PF algorithm to estimate $\psi$.\
However, being $\psi$ constant over time, it is meaningless to sample its values sequentially: the first extraction of $\psi$ at the initial stage 0 (i.e., $\psi_0^n$) generates a constant path of particles for the parameter ($\psi_t^n=\psi_0^n$). \footnote{To be more precise, at the resampling step, also $\psi^n$ is resampled, but the possible values that it may take are the $N$ ones obtained at the initial time 0 extraction.}\
Therefore, Liu and West (2001) proposed a modified particle filter, called \textit{Liu and West filter} (LWF) that allows to resample the parameter over time from a continuous distribution. In this way, at every time, the support of $\psi$ is not limited to the initially sampled $N$ values. \
Here, we describe the LWF that makes use of the normal distribution to sample sequentially the parameter. Moreover, consider the simpler version of the bootstrap filter.\footnote{Clearly, the LWF can be also used within the more general frameweorks of the GPF and the APF.}\
Let the transition kernel and the likelihood depend on the parameter $\psi$ (i.e., $P_t(x_{t-1},dx_t;\psi)$ and $f_t(y_t|x_t;\psi)$) and let $\pi(\cdot) $ be the prior distribution for the parameter $\psi$.\
The LWF can be described by the following algorithm.
\begin{enumerate}
    \item \textbf{Inital stage: }For $n=1,...,N$, at time 0 sample: $X_0^n\thicksim\Pd_0(dx_0)$ and $\psi_0^n\thicksim \pi(\psi_0)$.

@@

```{r cll_plot, echo=F, fig.align='center', fig.cap="Relative log-likelihood", fig.height=3, fig.pos='H', fig.width=10, results=F}
load("SV/cll_plot.Rda")
cll_plot+
  theme_bw()+
  scale_x_date(date_breaks = "3 months",date_minor_breaks="1 month",date_labels = "%m/%y")+
  labs(x="Time (month/year)",y="")
```
Overall, this analysis shows how a model with stochastic
volatility is better at describing and predicting the considered time-frame of S&P500 returns. As can be seen from the graph and perhaps unsurprisingly,
the density forecast given the stochastic volatility model represents
a particularly significant improvement in the period of extreme values of returns in March 2020.

>>>>

---
title: "Final draft"

@@

output:
  pdf_document: 
   latex_engine: pdflatex
   extra_dependencies: ["float"]
   number_sections: true

header-includes:
- \def\code#1{\texttt{#1}}
- \usepackage{amssymb}
- \usepackage{txfonts}
- \usepackage{amsmath}
- \usepackage[bindingoffset=1.5cm, left=3cm, right=3cm, top=3cm, bottom=3cm]{geometry}
- \usepackage{graphicx}
- \usepackage{MnSymbol} %allows to insert the QED at the end of the proofs
- \usepackage{hyperref} %set the hyperlink to cite sections, equations...
- \urlstyle{same}
- \usepackage{comment}
- \usepackage{enumitem}
- \usepackage{ntheorem}
- \usepackage{lipsum}
- \usepackage{xcolor}
- \theoremstyle{break}
- \newtheorem{thm}{Theorem}% theorem counter resets every \subsection
- \renewcommand{\thethm}{\arabic{thm}}
- \newtheorem{proposition}{Proposition}
- \newtheorem{assumption}{Assumption}
- \newtheorem{lemma}{Lemma}
- \newtheorem{definition}{Definition}
- \theoremstyle{nonumberplain}
- \newtheorem{proof*}{Proof.}
- \renewcommand{\P}{\mathbb{P}}
- \newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
- \def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
- \setcounter{MaxMatrixCols}{10}
- \setlength{\oddsidemargin}{.0in}
- \setlength{\textwidth}{6 in}
- \setlength{\topmargin}{-.1in}
- \setlength{\textheight}{8.1in}
- \renewcommand{\baselinestretch}{1.5}
- \newcommand{\E}{\mathbb E}
- \newcommand{\Q}{\mathbb Q}
- \newcommand{\M}{\mathbb M}
- \newcommand{\R}{\mathbb R}
- \newcommand{\B}{\mathcal B}
- \newcommand{\X}{\mathcal X}
- \newcommand{\Y}{\mathcal Y}
- \newcommand{\Pd}{\mathbb P}
- \newcommand{\N}{\mathcal N}
- \newtheorem{algorithm}{Algorithm}[section]
- \usepackage{float}
- \floatplacement{figure}{}

@@

Let $\varphi: \X \rightarrow \R$ be a measurable, integrable and bounded real-valued function, where $(\X,\B(\X),\Q)$ is a probability space.\footnote{$\B(\X)$ is the Borel sigma-algebra of $\X$.}
Suppose that the probability measure $\Q$ admits a density $q(x)$ with respect to some measure $\nu(dx)$ and it is feasible to sample from that density.\
To estimate the expected value of $\varphi$ with respect to the density $q$, the basic Monte Carlo (MC) method would require to sample $N$ times from the target density $q$ and to compute the average of the function $\varphi$ values. $\hat \varphi_{MC}$ is the standard MC estimator:
\begin{equation*}
   \hat \varphi_{MC}=\frac{1}{N} \sum_{n=1}^N\varphi (X^n)\approx \E_q(\varphi)~~~X^n\thicksim q(\cdot)
\end{equation*}
However, in many circumstances either it is impossible to sample from the target density or there exist another density - call it $m(\cdot)$ - different from $q$ that produces "more efficient" estimates. 
In all the cases for which sampling takes place from a distribution different from the "correct" (i.e., the true) one, importance sampling and resampling methods can be applied (as augmented versions of the standard MC algorithm).

@@

Let us consider a more general state-space model, where the state vector includes a time-constant parameter vector $\psi \in \Psi$, where $\Psi$ is the parameter space. Since we can interpret the parameter as a state with the law of motion $\psi_t=\psi_{t-1}=\psi$, then it is possible to adopt a PF algorithm to estimate $\psi$.\
However, being $\psi$ constant over time, it is meaningless to sample its values sequentially: the first extraction of $\psi$ at the initial stage 0 (i.e., $\psi_0^n$) generates a constant path of particles for the parameter ($\psi_t^n=\psi_0^n$). \footnote{To be more precise, at the resampling step, also $\psi^n$ is resampled, but the possible values that it may take are the $N$ ones obtained at the initial time 0 extraction.}\
Therefore, Liu and West (2001) proposed a modified particle filter, called \textit{Liu and West filter} (LWF) that allows to resample the parameter over time from a continuous distribution. In this way, at every time, the support of $\psi$ is not limited to the initially sampled $N$ values. \
Here, we describe the LWF that makes use of the normal distribution to sample sequentially the parameter. Moreover, consider the simpler version of the bootstrap filter.\footnote{Clearly, the LWF can be also used within the more general frameweorks of the GPF and the APF.}\
Let the transition kernel and the likelihood depend on the parameter $\psi$ (i.e., $P_t(x_{t-1},dx_t;\psi)$ and $f_t(y_t|x_t;\psi)$) and let $\pi(\cdot)$ be the prior distribution for the parameter $\psi$.\
The LWF can be described by the following algorithm.
\begin{enumerate}
    \item \textbf{Inital stage: }For $n=1,...,N$, at time 0 sample: $X_0^n\thicksim\Pd_0(dx_0)$ and $\psi_0^n\thicksim \pi(\psi_0)$.

@@

```{r cll_plot, eval=FALSE, fig.align='center', fig.cap="Relative log-likelihood", fig.height=3, fig.pos='H', fig.width=10, include=FALSE, results=F}
load("SV/cll_plot.Rda")
cll_plot+
  theme_bw()+
  scale_x_date(date_breaks = "3 months",date_minor_breaks="1 month",date_labels = "%m/%y")+
  labs(x="Time (month/year)",y="")
```
Overall, this analysis shows how a model with stochastic
volatility is better at describing and predicting the considered time-frame of S&P500 returns. As can be seen from the graph and perhaps unsurprisingly,
the density forecast given the stochastic volatility model represents
a particularly significant improvement in the period of extreme values of returns in March 2020.
